Tail associations and their impact on extinction risk

### 2019-10-30 MainText.Rmd
edits to results

Because the deterministic, one-patch versions of our models showed undercompensatory
dynamics for $r<r_c$ (and therefore for $r$ on the left end of the range $(r_{min},r_{bf})$),
and showed overcompensatory dynamics for $r>r_c$, Figs \ref{fig_extrisk_allmodel_localdisp}
and \ref{SM-fig_extrisk_allmodel_globaldisp} suggest the hypothesis that the
opposite effects of tail association on extinction risk correspond to the domains
in which models exhibit under- versus overcompensatory dynamics. This was 
approximately, but not exactly, true. Careful inspection of Figs 
\ref{SM-fig_ricker_risk_local_disp_scl_1} - \ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}
and Table

<!--Q3 : metapopulation results : switching near rc for d=0-->
Following Figs. (\ref{SM-fig_ricker_risk_local_disp_scl_1} - 
\ref{SM-fig_risk_global_disp_abrewer_r_vary_scl_0.6}) one can reasonably hypothesize
that the switching between the extinction risk 
generated by left-tail associated and right-tail associated noise 
occurred near the critical parameter $r_c$. For simplicity, we only considered 
$d=0$ case, i.e., when there was no dispersal between patches. 
Fig. \ref{fig_extrisk_d0} demonstrates higher extinction risk 
with left-tail associated noise for $r<r_c$ and with right-tail associated noise 
for $r>r_c$. $r_c$ was the vertical line in Fig. \ref{fig_extrisk_d0} 
indicating the value of $r$ for each model where their deterministic analog switches 
from under- to over-compensatory dynamics. Though the switching visually appeared far away
from the theoretical $r_c$ value for Verhulst and Austin-Brewer model, they were 
consistent with our hypothesis as they had comparatively a smaller range of $r$ than the other models.

@@

<!--Summary tables for 6 nonlinear models used-->
```{r read_tab_model_summary,echo=F, results='asis',message=F}
library(kableExtra)
library(dplyr)
dt<-readRDS("./Results/tab_model_summary.RDS")
knitr::kable(dt, escape=F,
             format="latex",digits=3, align="c",table.env='table*',
             caption = "Summary of the six density-dependent stochastic metapopulation models we used. The population growth rate in habitat patch $i$ at time $t$, $\\lambda_i(t)$ (second column), is $P_i(t+1)/P_i(t)$, and defines how the model works within each patch. Here $P_i(t)$ is the population in patch $i$ at time $t$. The growth rate depends on model parameters, $P_i(t)$, and the environment $\\epsilon_i(t)$ in patch $i$ at time $t$. $\\epsilon_i(t)$ was normally distributed with mean $0$ and standard deviation $\\sigma$, and values were generated independently across times. $P_e$ is the equilibrium of the deterministic, one-patch version of the model. $r_{min}$ is the smallest $r$, given fixed values of the other parameters, for which $P_e$ was non-negative. $r_c$ is the value of $r$ for which a transition occurs from under- to overcompensatory dynamics (see main text). $r_{bf}$ is the value of $r$ for which a transition occurs from a stable to an unstable equilbrium (a bifurcation; see main text). We used $r$ such that $r_{min}<r<r_{bf}$. Dispersal between patches was also implemented (see main text).\\label{tab_model_summary}",
             booktabs=T, linesep = "\\addlinespace"
             )%>%column_spec(c(1,3,5,6,7),width="1.4 cm")%>%kable_styling(latex_options = "basic",position = "center")%>%column_spec(2, width = "3.8 cm")
```
<!--***DAN: Shyamolina, can you please make the header of the second column read simply \lambda_i(t)? Also, can you add a column for r_min, please?
***Shya: Done, I added r_min value, if a population starts with exactly r=r_min, then P_e remains at 0, it should start with r>r_min to get a positive equilibrium.-->

>>>>

Because the deterministic, one-patch versions of our models showed undercompensatory
dynamics for $r<r_c$ (and therefore for $r$ on the left end of the range $(r_{min},r_{bf})$),
and showed overcompensatory dynamics for $r>r_c$, Figs \ref{fig_extrisk_allmodel_localdisp}
and \ref{SM-fig_extrisk_allmodel_globaldisp} suggest the hypothesis that the
opposite effects of tail association on extinction risk correspond to the domains
in which models exhibit under- versus overcompensatory dynamics. This was 
approximately, but not exactly, true. Careful inspection of Figs 
\ref{fig_risk_local_disp_verhulst_r_vary_scl_0.6} - \ref{fig_risk_local_disp_abrewer_r_vary_scl_0.6} and Figs \ref{fig_risk_global_disp_verhulst_r_vary_scl_0.6} - \ref{fig_risk_global_disp_abrewer_r_vary_scl_0.6}  
and the $r_c$ values of 
Table \ref{tab_model_summary} reveals that, for the Velhurst and Austin-Brewer models,
the switch occurs at a value of $r$ less than $r_c$ for these models. For the example
case of $d=0$, plots of extinction risk versus $r$ (Fig. \ref{fig_extrisk_d0}) seem to indicate that,
for the Ricker, Hassell, and Pennycuick models, left-tail association accentuates and
left-tail association mitigates extinction risk for undercompensatory dynamics ($r<r_c$),
whereas the reverse occurs for overcompensatory dynamics ($r>r_c$). Fig. \ref{fig_extrisk_d0}
reveals a similar reversal for the Maynard Smith, Velhurst and Austin-Brewer models; but 
it occurs at a value of $r$ close to but distinguishable from $r_c$.

<!--***DAN: stopped here, also need to do the caption of the last figure-->

@@

<!--Summary tables for 6 nonlinear models used-->
```{r read_tab_model_summary,echo=F, results='asis',message=F}
library(kableExtra)
library(dplyr)
dt<-readRDS("./Results/tab_model_summary.RDS")
knitr::kable(dt, escape=F,
             format="latex",digits=3, align="c",table.env='table*',
             caption = "Summary of the six density-dependent stochastic metapopulation models we used. The population growth rate in habitat patch $i$ at time $t$, $\\lambda_i(t)$ (second column), is $P_i(t+1)/P_i(t)$, and defines how the model works within each patch. Here $P_i(t)$ is the population in patch $i$ at time $t$. The growth rate depends on model parameters, $P_i(t)$, and the environment $\\epsilon_i(t)$ in patch $i$ at time $t$. $\\epsilon_i(t)$ was normally distributed with mean $0$ and standard deviation $\\sigma$, and values were generated independently across times. $P_e$ is the equilibrium of the deterministic, one-patch version of the model. $r_{min}$ is the smallest $r$, given fixed values of the other parameters, for which $P_e$ was non-negative. $r_c$ is the value of $r$ for which a transition occurs from under- to overcompensatory dynamics (see main text). $r_{bf}$ is the value of $r$ for which a transition occurs from a stable to an unstable equilbrium (a bifurcation; see main text). We used $r$ such that $r_{min}<r<r_{bf}$. Dispersal between patches was also implemented (see main text). Formulas are given for $P_e$ but specific values which pertain for the listed parameters are given for $r_{min}$, $r_c$ and $r_{bf}$. See Appendix \\ref{SM-model_stability} for the formulas that gave these values.\\label{tab_model_summary}",
             booktabs=T, linesep = "\\addlinespace"
             )%>%column_spec(c(1,3,5,6,7),width="1.4 cm")%>%kable_styling(latex_options = "basic",position = "center")%>%column_spec(2, width = "3.8 cm")
```
